---
title: "Data"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(readr)
library(dplyr)
library(lubridate)
library(rvest)
library(gganimate)
library(gifski)
library(png)
```

This graph shows that total number of deaths from the years 1982 to 2019. The graph shows the total deaths for the ages 16,18,21 and 30 and above. Each graph also shows how much of the total deaths are caused by having a blood alchohol concentration(bac) of 0.08 or higher, otherwise known as drinking and driving.

```{r, echo=FALSE}
website<-read_html("https://www.iihs.org/topics/fatality-statistics/detail/teenagers#alcohol-involvement")

alldata <-website%>% html_elements("#alcohol-involvement .is-fullwidth:nth-child(4) th , #alcohol-involvement .is-fullwidth:nth-child(4) td") %>% html_text2()
years<-alldata[seq(27,520,13)]

driverskilled_16<-alldata[seq(28,520,13)]
bac_16<-alldata[seq(29,520,13)]
percent_16<-alldata[seq(30,520,13)]

driverskilled_18<-alldata[seq(31,520,13)]
bac_18<-alldata[seq(32,520,13)]
percent_18<-alldata[seq(33,520,13)]

driverskilled_21<-alldata[seq(34,520,13)]
bac_21<-alldata[seq(35,520,13)]
percent_21<-alldata[seq(36,520,13)]

driverskilled_30<-alldata[seq(37,520,13)]
bac_30<-alldata[seq(38,520,13)]
percent_30<-alldata[seq(39,520,13)]

graph1<-tibble(years,driverskilled_16,driverskilled_18,driverskilled_21,driverskilled_30,bac_16,
              bac_18,bac_21,bac_30) %>%  
  mutate(across(everything(),~str_replace_all(.,",",""))) %>% 
  mutate(across(everything(),as.numeric)) %>% 
  pivot_longer(cols=-years,
               names_sep = "_",names_to=c("type","age"),values_to="totalvalue") %>% 
 ggplot(aes(x=years,y=totalvalue,fill=type)) + geom_col(position="identity") + 
  transition_time(as.integer(age))+
  labs(title="Total Motor Vehicle Deaths from 1982-2019, transitioning through ages 16-30",
       x="Year",y="Total Deaths", caption="Source: iihs.org", 
       subtitle="Through the ages of 16-30, the amount of deaths increased, but throughout the years, the total deaths decreased, except for age 30") + 
  theme_minimal() + 
  theme(panel.spacing = unit(2, "lines")) + 
  scale_fill_discrete(name = "Cause",   labels = c("Alcohol","Other")) + 
  labs(subtitle = "Age: {frame_time}")

graph1
```

Taking a closer look at the percentages, as age increases, the percentage of alcohol related deaths increase, then decrease. But the overall trend is positive, which means that alcohol is a more prevalent cause in vehicle deaths.   

```{r, echo = FALSE}
graph2<-tibble(years,percent_16,percent_18,percent_21,percent_30) %>% 
  mutate(across(everything(),~str_replace_all(.,",",""))) %>% 
  mutate(across(everything(),as.numeric)) %>%
  pivot_longer(cols=-years,names_sep="_", names_to = c(NA,"age"),values_to = "percent") %>% 
  ggplot(aes(years,percent))+geom_col(color="steelblue") + theme_minimal() +  
  theme(panel.spacing = unit(2, "lines")) + 
  transition_time(as.integer(age)) + 
  labs(subtitle = "Year: {frame_time}") + 
  labs(title="Percent of Total Deaths that are Alcohol Related from ages 16-30, from 1982-2019", subtitle="Around 21 years of age is when the percentage is the highest through all the years",x="Year",y="Percent of Total Vehicle Deaths thatare alcohol related (BAC >= 0.08)")

graph2
```
